###Title: "Practical Machine Learning - Final Project"
--Author: "Viet Nguyen"
--Date: "August 12, 2016"
--output: html_document


## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


## Preparation step: set WD and load all libraries + data
- Working Directory and load data into R 
```{r}
setwd("C:/Users/user/Desktop/RWD/PML")
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(corrplot)
library(knitr)  

```

- Loading Data
```{r}
set.seed(12345)

trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!","")) 
```

- Partioning the training set
```{r}
inTrain <- createDataPartition(training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```
## Cleaning the data

- Remove NearZeroVariance variables
```{r}
nzv <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[,nzv$nzv==FALSE]

nzv<- nearZeroVar(myTesting,saveMetrics=TRUE)
myTesting <- myTesting[,nzv$nzv==FALSE]
myTraining <- myTraining[c(-1)]

```

- Clean variables with more than 60% NA
```{r}
trainingV3 <- myTraining
for(i in 1:length(myTraining)) {
    if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .7) {
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTraining[i]), names(trainingV3)[j]) ) == 1)  {
                trainingV3 <- trainingV3[ , -j]
            }   
        } 
    }
}

# Set back to the original variable name
myTraining <- trainingV3
rm(trainingV3)

```

- Transform the myTesting and testing data sets
```{r}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58])  # remove the classe column
myTesting <- myTesting[clean1]         # allow only variables in myTesting that are also in myTraining
testing <- testing[clean2]             # allow only variables in testing that are also in myTraining

dim(myTesting)
```


- Coerce the data into the same type
```{r}
for (i in 1:length(testing) ) {
    for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) == 1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}

# To get the same class between testing and myTraining
testing <- rbind(myTraining[2, -58] , testing)
testing <- testing[-1,]
```

###Prediction with Decision Trees
```{r}

set.seed(12345)
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
#predictionsA1
#install.packages('e1071', dependencies=TRUE)
cmtree <- confusionMatrix(predictionsA1, myTesting$classe)
plot(cmtree$table, col = cmtree$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy =", round(cmtree$overall['Accuracy'], 4)))
fancyRpartPlot(modFitA1)

```


###Prediction with Random Forests
```{r}
set.seed(12345)
modFitB1 <- randomForest(classe ~ ., data=myTraining)
predictionB1 <- predict(modFitB1, myTesting, type = "class")
cmrf <- confusionMatrix(predictionB1, myTesting$classe)
plot(modFitB1)
plot(cmrf$table, col = cmtree$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(cmrf$overall['Accuracy'], 4)))
```


###Prediction with Generalized Boosted Regression
```{r}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
#modFitGBM  <- train(classe ~ ., data=myTraining, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
#modFitGBM$finalModel
#install.packages('C50', repos='cran.us.r-project.org')
#fitControl <- trainControl(method = "repeatedcv",number = 5,repeats = 1)
#gbmFit1 <- train(classe ~ ., data=myTraining, method = "gbm",trControl = fitControl, verbose = FALSE)
#gbmFinMod1 <- gbmFit1$finalModel
#gbmPredTest <- predict(gbmFit1, newdata=myTesting)
#gbmAccuracyTest <- confusionMatrix(gbmPredTest, myTesting$classe)
#plot(gbmFit1, ylim=c(0.9, 1))
```


###Predicting Results on the Test Data
Random Forests gave an Accuracy in the myTesting dataset of 99.89%, which was more accurate that what I got from the Decision Trees or GBM. The expected out-of-sample error is 100-99.89 = 0.11%.




